---
title: "Food at College"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, warning=FALSE, message=FALSE}
# From your file menu (at the top of R-Studio) select:
# "Session -> Set working directory -> To source file location"
# Then play this chunk to get the data into R.
library(mosaic)
library(car)
library(DT)
library(pander)
library(readr)
library(plotly)
food <- read_csv("../../Data/food.csv") #food.csv is in the Data folder...
food$Grade_Level <- case_when(food$grade_level==1 ~ 'Freshmen',
                             food$grade_level==2 ~ 'Sophomores',
                             food$grade_level==3 ~ 'Juniors',
                             food$grade_level==4 ~ 'Seniors')
```


----

<!-- Instructions:

 Use the Food dataset and a Kruskal-Wallis Test(s) to answer an interesting question(s) that you come up with.
 
 About the dataset:
 
  * GPA - numerical, self-reported current college GPA

  
  * weight - numerical, self-reported weight in lbs


  * gender - "Female", "Male"


  * breakfast - "Cereal", "Donut" 
      The participants were shown a picture of "Cereal" and a "Donut" and asked which one of these pictures they associate with the word "breakfast."


  * calories_chicken - guessing calories in chicken piadina from the options of:
     1 - 265 
     2 - 430 
     3 - 610 
     4 - 720
    (the variable shows the actual number of calories participants selected)


   * calories_day - Importance of consuming calories per day
     1 - i dont know how many calories i should consume 
     2 - it is not at all important 
     3 - it is moderately important 
     4 - it is very important


   * comfort_food - List 3-5 comfort foods that come to mind. 
     Open ended 


   * comfort_food_reasons - What are some of the reasons that make you eat comfort food?      (i.e., anger, sadness, happiness, boredom, etc) - list up to three
     Open ended


   * comfort_food_reasons_coded
     1 - stress
     2 - boredom
     3 - depression/sadness
     4 - hunger
     5 - laziness
     6 - cold weather
     7 - happiness 
     8 - watching tv
     9 - none 


    * cook - how often do you cook?
     1 - Every day 
     2 - A couple of times a week 
     3 - Whenever I can, but that is not very often  
     4 - I only help a little during holidays 
     5 - Never, I really do not know my way around a kitchen


    * cuisine - what type of cuisine did you eat growing up?
     1 - American
     2 - Mexican.Spanish
     3 - Korean/Asian
     4 - Indian
     5 - American inspired international dishes
     6 - other


    * diet_current - describe your current diet
     open ended

	
    * diet_current_coded
     1 - healthy/balanced/moderated/
     2 - unhealthy/cheap/too much/random/
     3 - the same thing over and over
     4 - unclear


    * drink - which picture do you associate with the word “drink”?
     1 - orange juice
     2 - soda 


    * eating_changes  - Describe your eating changes since the moment you got into college?
     Open ended 


    * eating_changes_coded
     1 - worse
     2 - better
     3 - the same
     4 - unclear


    * eating_changes_coded1
     1 - eat faster
     2 - bigger quantity
     3 - worse quality 
     4 - same food
     5 - healthier
     6 - unclear
     7 - drink coffee 
     8 - less food
     9 - more sweets
     10 - timing 
     11 - more carbs or snacking
     12 - drink more water
     13 - more variety


    * eating_out - frequency of eating out in a typical week 
     1 - Never 
     2 - 1-2 times 
     3 - 2-3 times 
     4 - 3-5 times 
     5 - every day


    * employment - do you work? 
     1 - yes full time 
     2 - yes part time 
     3 - no
     4  - other


    * ethnic_food - How likely to eat ethnic food 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * exercise - how often do you exercise in a regular week?
     1 - Everyday 
     2 - Twice or three times per week 
     3 - Once a week
     4 - Sometimes 
     5 - Never


    * father_education - 
     1 - less than high school 
     2 - high school degree 
     3 - some college degree 
     4 - college degree 
     5 - graduate degree 


    * father_profession - what is your father profession?
     Open ended


    * fav_cuisine - What is your favorite cuisine?
     Open ended


    * fav_cuisine_coded
     0-none
     1 - Italian/French/greek
     2 - Spanish/mexican
     3 - Arabic/Turkish
     4 - asian/chineses/thai/nepal
     5 - American
     6 - African 
     7 - Jamaican
     8 - indian


    * fav_food - was your favorite food cooked at home or store bought? 
     1 - cooked at home 
     2 - store bought 
     3 - both bought at store and cooked at home


    * food_childhood - what was your favorite childhood food?
     Open ended



    * fries - which of these pictures you associate with word fries? 
     1 - Mcdonald’s fries
     2 - home fries

 
    * fruit_day - How likely to eat fruit in a regular day 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * grade_level – 
     1 - freshman 
     2 -Sophomore 
     3 - Junior 
     4 - Senior
     
      
    * greek_food - How likely to eat greek food when available?
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 
     
      
    * healthy_feel – how likely are you to agree with the following statement: “I feel very healthy!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
    * healthy_meal – what is a healthy meal? Describe in 2-3 sentences.
      Open ended
      
      
    * ideal_diet – describe your ideal diet in 2-3 sentences
      Open ended
      
      
    * Ideal_diet_coded
      1 – portion control
      2 – adding veggies/eating healthier food/adding fruit	
      3 – balance
      4 – less sugar
      5 – home cooked/organic
      6 – current diet
      7 – more protein
      8 – unclear
      
      
     * income
      1 - less than $15,000 
      2 - $15,001 to $30,000 
      3 - $30,001 to $50,000 
      4 - $50,001 to $70,000 
      5 - $70,001 to $100,000 
      6 - higher than $100,000
      
     * indian_food – how likely are you to eat indian food when available
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * Italian_food – how likely are you to eat Italian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * life_rewarding – how likely are you to agree with the following statement: “I feel life is very rewarding!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
     * marital_status
      1 -Single 
      2 - In a relationship 
      3 - Cohabiting 
      4 - Married 
      5 - Divorced 
      6 - Widowed
      
      
     * meals_dinner_friend – What would you serve to a friend for dinner?
      Open ended
      
      
     * mothers_education 
      1 - less than high school 
      2 - high school degree 
      3 - some college degree 
      4 - college degree 
      5 - graduate degree
      
      
     * mothers_profession – what is your mother’s profession? 
      
      
     * nutritional_check - checking nutritional values frequency 
      1 - never 
      2 - on certain products only 
      3 - very rarely 
      4 - on most products 
      5 - on everything
      
      
     * on_off_campus – living situation
      1 - On campus 
      2 - Rent out of campus 
      3 - Live with my parents and commute 
      4 - Own my own house
      
      
     * parents_cook - Approximately how many days a week did your parents cook? 
      1 - Almost everyday 
      2 - 2-3 times a week 
      3 - 1-2 times a week 
      4 - on holidays only 
      5 - never
      
      
     * pay_meal_out - How much would you pay for meal out? 
      1 - up to $5.00 
      2 - $5.01 to $10.00 
      3 - $10.01 to $20.00 
      4 - $20.01 to $30.00 
      5 - $30.01 to $40.00 
      6 - more than $40.01 
      
      
     * Persian_food - How likely to eat persian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * self_perception_weight - self perception of weight 
      6 - i dont think myself in these terms 
      5 - overweight 
      4 - slightly overweight 
      3 - just right 
      2 - very fit 
      1 - slim 
      
      
     * Which of the two pictures you associate with the word soup?
      1 – veggie soup
      2 – creamy soup
      
      
     * sports - sports – do you do any sporting activity?
      1 - Yes 
      2 - No 
      99 – no answer
      
      
     * thai_food - How likely to eat thai food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * tortilla_calories - guessing calories in a burrito sandwhich from Chipolte?
      1 - 580 
      2 - 725 
      3 - 940 
      4 - 1165
      
     * turkey_calories - Can you guess how many calories are in the foods shown below?
     (Panera Bread Roasted Turkey and Avocado BLT)
      1 - 345 
      2 - 500 
      3 - 690 
      4 - 850
      
      
     * type_sports – what type of sports are you involved?
      Open-ended
      
     * veggies_day - How likely to eat veggies in a day? 
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4- likely 
      5 - very likely
      
     * vitamins – do you take any supplements or vitamins?
      1 – yes
      2 – no
      
     * waffle_calories - guessing calories in waffle potato sandwhich 
      1 - 575 
      2 - 760 
      3 - 900 
      4 - 1315


Note this dataset and description come from: © 2020 Kaggle Inc for further details visit:

[Food choices on Kaggle](https://www.kaggle.com/borapajo/food-choices?select=food_coded.csv)


If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->
**Comments to Critiquers:** 

<!--End your comments to your critiquers before this line.-->

----

<!--Begin writing your Analysis after this point.-->

## Background

I want to see whether students give a rating closer to agree or strongly agree for the "I feel very healthy!" statement when they have more years in college. My personal assumption is that the more time you spend in college, you become a stronger critical thinker and reflect more on what it is that you do within the aspects of your life. I also think that you eat out less because you are more concerned with saving money for when you finish school. Hence, I believe that when you are closer to leaving college, the average rating will be closer to an agree or strongly agree rating for the "I feel very healthy!" statement.

## Introduction

I want to conduct a Kruskal-Wallis test to see if the junior and seniors see themselves as more healthy than freshman and sophomores. As previously mentioned, I will look at ratings for the statement: "I feel very healthy!" A 1 is strongly agree and a 10 is strongly disagree. I am assuming students in their later years of college will have a rating closer to 1 (strongly agree).

## Questions and Hypotheses

The question I am trying to answer is: 

Is the average rating for the statement "I feel very healthy!" different for at least one college classification group (freshmen, sophomores, juniors, seniors)? 

$$
  H_0: \text{Mean ratings for all college groups are the same}
$$

$$
  H_a: \text{At least one mean rating for one college group is different than the others}
$$

Significance level will be: α=0.05

## Data Analysis

```{r message=FALSE, warning=FALSE}
student.aov <- aov(healthy_feeling ~ Grade_Level, data=food)
#summary(student.aov) %>% 
#pander()
plot(student.aov, which=1, pch=16)

```

As you can see, my plot for variance is constant. This means that it satisfies one of the requirements to perform an ANOVA test.

```{r message=FALSE, warning=FALSE}
qqPlot(student.aov$residuals, id=FALSE, main='QQPlot')

```

Some data points in the QQPlot are outside of the boundary lines, which means the QQPlot doesn't show normality of the residuals. This means I have to perform a Kruskal-Wallis test.

```{r message=FALSE, warning=FALSE}
kruskal.test(healthy_feeling ~ Grade_Level, data = food) %>% pander()

``` 

The Kruskal-Wallis test shows the P-value is 0.5172. Therefore, we have insufficient evidence to reject the null hypothesis. There is no significant difference in the average healthy feeling rating for at least one group (freshmen, sophomores, juniors, and seniors groups).

## Graphical Summary
```{r message=FALSE, warning=FALSE}
plot_ly(food, y=~healthy_feeling, x=~Grade_Level, type='box', fillcolor='red', line=list(color='black', width=3), marker = list(color='orange', line = list(color='red', width=1))) %>%
layout(title='Healthy Feeling Rating by College Classification \n (Freshmen, Sophomores, Juniors, Seniors)', xaxis=list(title='College Classification'), yaxis=list(title='Healthy Feeling Rating'))

#xyplot(healthy_feeling ~ as.factor(GradeLevel), data=food, main='Healthy Feeling Rating by College Classification \n (Freshman, Sophomores, Juniors, Seniors)', xlab='College Classification', ylab='Healthy Feeling', type=c("p","a"), col='red')

```

The boxplot shows the minimum, Q1, median, Q3, and Max healthy feeling rating for the freshmen, sophomores, juniors, and seniors. As we can see, the median healthy feeling rating is very similar for all groups. For freshmen and juniors, the median healthy feeling rating is 6, for sophomores and seniors it is 5. 

## Numerical Summary

#### Summary and Sample Size for Freshmen , Sophomores, Juniors, and Seniors

```{r message=FALSE, warning=FALSE}
food %>% 
   group_by(Grade_Level) %>% 
   summarise(Mean_Rating = mean(healthy_feeling),
             Median_Rating = median(healthy_feeling),
             sd_Rating = sd(healthy_feeling), 
             Sample_Size = n()) %>%
  pander()

```

My numerical summary shows the median, mean, and standard deviation for the healthy feeling rating for each group (freshmen, sophomores, juniors, and seniors). It also shows the sample size for each group. The standard deviation for each group's healthy feeling rating is very close to the same. The average healthy feeling rating for each group is very close as well. For the average healthy feeling rating most groups (freshmen, sophomores, and seniors) were at about 5. Only juniors had an average healthy feeling rating at 6 (specifically 6.071). The median healthy feeling rating for freshmen and juniors was 6 and 5 for seniors and sophomores. The median for each grade level group is close to the same to their averages.

## Interpretation

From the data, we can conclude that the average healthy feeling rating by freshmen, sophomores, juniors, and seniors is about the same. We failed to reject the null hypothesis and our data shows how close the healthy feeling ratings are for each group. There was no statistically significant mean difference from at least one of the groups. Therefore, the assumption I made at the beginning of the test was proven false. 







