---
title: "Simple Linear Regression"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
# Load your libraries
library(car)
library(tidyverse)
library(pander)
library(mosaic)
library(DT)
# Load your data after saving a csv file in your Data folder.
# You can use either 
#   someName <- read.csv("../../Data/YourDataFileName.csv", header=TRUE)
# or
library(readr)
DataforLinearRegression <- read_csv('../../Data/DataforLinearRegression.csv')

#Don't forget to run "Session -> Set Working Directory -> To Source file location"

```


<!-- Directions

Come up with a simple linear regression of your own creation. 

* Write the Background and clearly state your question. Then collect data and perform the linear regression analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.

* You should collect both the quantitative response variable and the quantitative explanatory variable yourself. This can be data you collect from 
    - the internet (like weather data, Facebook data, or sports data) 
    - your own experiment 
    - your own observational study
* However,     
    - The data should not be a single pre-existing dataset. It must be something that you piece together yourself.

* You need at least $n=3$ $(X,Y)$ pairs to perform the regression. However, your regression will only be useful if you can get around $n=10$ or more observations.

-->
## House Prices in Denver, Colorado

### Introduction/Background

Colorado is a beautiful state, but are the prices of homes beautiful as well? Within the last couple of years, prices for homes have skyrocketed. Because of that, I have decided to conduct a linear regression analysis that compares home prices and square footage for houses in Denver, Colorado (with a sample of 30 houses). I gathered my information from a website called Zillow which showed me the houses for sale in Denver, Colorado along with the details of the house such as the square footage. I gathered a sample of 30 different homes from random communities.

### Question/Hypothesis

The question I am trying to answer is: How strong of a relationship is there between the price of a home and the square footage in Denver, Colorado?

The hypothesis below states that the slope is equal to zero. The alternative hypothesis states that the slope is not equal to zero. 
$$
  H_0: \beta_1 = 0
$$
$$
  H_a: \beta_1 \neq 0
$$
Here is the significance level for this analysis: α=0.05

### Validation of the Model Appropriateness

```{r, message=FALSE, warning=FALSE}
DenverH.lm <- lm(Home_Price ~ Square_Footage, data = DataforLinearRegression)
#par(mfrow=c(1,3))
```

```{r, message=FALSE, warning=FALSE}
plot(DenverH.lm, which=1)
```

The residuals vs fitted-values plot checks to see if there is a linear relationship and if there is a constant variance. There seems to be no pattern at all. This means there is a linear relationship and a constant variance. 

```{r, message=FALSE, warning=FALSE}
qqPlot(DenverH.lm$residuals, id=FALSE)
```

The normality of the error terms can be checked through this QQPlot. It seems that most points are within the dashed boundary lines. There is only two points that are not. However, since it is only two points outside of the dashed boundary lines, the error terms are considered to be normal. 

```{r, message=FALSE, warning=FALSE}
plot(DenverH.lm$residuals, main="Residuals vs Order")
```

There seems to be a general trend. This means that the independence of the error terms is questionable.

### Linear Regression

```{r, message=FALSE, warning=FALSE}
plot(Home_Price ~ Square_Footage, data=DataforLinearRegression, pch=21, bg='orange', col='black', 
     xlab='Square Footage of House', ylab='Price of the House (In thousands)', 
     main='The Price and Square Footage \n of a House in Denver, CO', las=1)

DenverH.lm <- lm(Home_Price ~ Square_Footage, data=DataforLinearRegression)

abline(DenverH.lm, lwd=3, col=rgb(.4,.4,.4,.2))
abline(v=seq(40,120,20), h=seq(40,120,20), lty=2, col=rgb(.6,.6,.6,.2))

```

### Analysis Details

The equation of the fitted line shown in the plot above is:

$$
  \underbrace{\hat{Y}_i}_\text{Home Price} = \overbrace{b_0}^\text{est. y-int} + \overbrace{b_1}^\text{est. slope} \underbrace{X_i}_\text{Square Footage}
$$

$$
  \text{Home Price} = 228,913.8 + 158.9 \times \text{(Square Footage)}
$$

This is saying that for every additional square foot, the average price of a home in Denver,Co will go up by $158.9, essentially $159. This is the change in the average Y-value. The Y-intercept tells us the average Y-value when X is zero. In this case, when there is no square footage, the average price of a house in Denver, Co is $228,913.8. In other words, it is telling us the average price of a plot of land without a home, the average cost of the land being $228,913.8. 

### Summary of the Regression

```{r, message=FALSE, warning=FALSE}
summary(DenverH.lm) %>% pander()
```

The summary of the regression analyzes price in thousands. It tells us that the slope is significant. The slope is 0.001052. This means that the relationship is meaningful. Our $R^2$ tells us that the regression model above accounts for 32.3% of the variance. The Adjusted $R^2$ is close to the same at 29.88%. This means that our data is not really close to the fitted regression line. 

### Numerical Summary

```{r, message=FALSE, warning=FALSE}
cor(DataforLinearRegression$Home_Price, DataforLinearRegression$Square_Footage) %>% 
  pander()
 

```

My numerical summary above shows the correlation between the price of a home in Denver, Co and the square footage. It appears we have a positive correlation. The strength of the two quantitative variables is not really strong or really week. It is around the middle in strength. 

### The Data

Here is the data of the experiment.

```{r, message=FALSE, warning=FALSE}
DataforLinearRegression1<-select(DataforLinearRegression, c('Square_Footage','Price'))

datatable(DataforLinearRegression1, options=list(lengthMenu = c(10,10,10)), extensions="Responsive")

```

### Conclusion

Based on the data, we have sufficient evidence to conclude that the slope is not equal to zero. The relationship is meaningful between the price of a house in Denver and the square footage. The data can be fairly accurate in predicting the price of a home in Denver based off of the square footage you want. For a recent college graduate, Denver probably is not the best place to move to because of how expensive it is. 



